---
title: "LiDAR Boulder Project"
format: html
editor: visual
---

## Goals

This is a personal project trying to find new bouldering zones in Squamish, BC. I am using publicly available LiDAR datasets (LidarBC) to create high resolution Digital Terrain Models (DTMs). 3D renders of these DTMs allow for terrain below the forest canopy to be studied and find hidden boulders!

In order to both create the DTMs and vizualize them I rely on the lidR package in R studio.

## What it looks like

This may take a while to get running because there's billions of points in this point cloud...

```{r, echo=FALSE,warning=FALSE,message=FALSE}

library(lidR)

apronbloc <- readLAS("C:/Users/lwill01.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/data/squam_laz/apron_laz.laz")
apron_dtm <- rasterize_terrain(apronbloc, 0.3, tin())

```

The first step is creating a DTM so we can see whats going on at the surface below the forest canopy.

Here is my DTM in 2D. You can see the areas of increased elevation where the bigger boulders are.

```{r,warning=FALSE}
plot(apron_dtm)
```

It is way more fun, and easier to see whats going on in 3D. The biggest boulder here is the Cacodemon boulder in Squamish with loads of famous routes on it.

```{r,echo=FALSE,message=FALSE,warning=FALSE}



library(rgl)
library(htmltools)

plot_dtm3d(apron_dtm)

htmlwidgets::saveWidget(
  rglwidget(),
  "apron_dtm3d.html",
  selfcontained = FALSE
)

rgl.close()


```

```{=html}
<iframe src="apron_dtm3d.html"
        width="100%"
        height="600"
        style="border:0;">
</iframe>
```

The same workflow can be replicated anywhere as long as the LiDAR data exists.

I repeated the workflow with a zone in Quebec I am very familiar with known as 'The Crown'.

Once again, I begin with a DTM.

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(raster)
crown <- readLAS("C:/Users/lwill01.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/data/squam_laz/crown_laz_clip.laz")
# Estimated rectangle from your red outline (adjust if needed)

crown_dtm <- rasterize_terrain(crown, 0.2, tin())


```

```{r}
plot(crown_dtm)
```

And can now visualize that DTM in 3D with the lidR package.

plot_dtm3d(crown_dtm)

```{r,echo=FALSE,message=FALSE,warning=FALSE}

library(rgl)
library(htmltools)

plot_dtm3d(crown_dtm)

htmlwidgets::saveWidget(
  rglwidget(),
  "crown_dtm3d.html",
  selfcontained = FALSE
)

rgl.close()
```

``` html
<iframe src="crown_dtm3d.html"
        width="100%"
        height="600"
        style="border:0;">
</iframe>
```

This zone displays some of the limitations of the DTM for finding boulders. The proximity of the climbing area to houses makes it difficult to delineate differences between the two. Consequently, some knowledge of the area will be useful in identifying new climbing zones, especially in more built areas.
