---
title: "LiDAR Boulder Project"
format: html
editor: visual
---

## Goals

This is a personal project trying to find new bouldering zones in Squamish, BC. I am using publicly available LiDAR datasets (LidarBC) to create high resolution Digital Terrain Models (DTMs). 3D renders of these DTMs allow for terrain below the forest canopy to be studied and find hidden boulders!

In order to both create the DTMs and vizualize them I rely on the lidR package in R studio.

## What it looks like

This may take a while to get running because there's billions of points in this point cloud...

```{r}
#|echo: false
library(lidR)

apronbloc <- readLAS("C:/Users/lwill01.stu/OneDrive - UBC/Documents/GitHub/E-Portfolio/data/squam_laz/apron_laz.laz")
apron_dtm <- rasterize_terrain(apronbloc, 0.3, tin())

```

Once we rasterize the terrain we can plot out DTM in 2D.

```{r}
plot(apron_dtm)
```

The next snippet of code lets us pan around the 3D render of the DTM and look boulders beneath the forest canopy.

```{r}
plot_dtm3d(apron_dtm)
```
